Triangulating iconicity: Coding analysis

### 2024-04-03 paper_02_main_analyses.Rmd
rename things and clean up

---
title: "Triangulating iconicity: Coding analysis"
author: "Mark Dingemanse & Stella Punselie"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

@@

# get consensus coding data
d = read_excel("data\\ideophones_coded.xlsx") %>% arrange(filename)

@@

d.scores = read_excel("data\\ideophones_guessability.xlsx") %>%
  dplyr::select(-category)
d <- left_join(d,d.scores,by=c("ideophone","language","study" = "paper"))

@@

# get ratings data
d.ratings <- read_xlsx("data\\ideophones_rated_means.xlsx") %>%
  dplyr::select(-category,-list,-item)
d <- left_join(d,d.ratings,by=c("filename","study","language"))

@@

@@

d %>%
  group_by(study) %>%
  summarize(max=max(score),min=min(score),mean=mean.na(score))

@@

ggsave("figures\\paper-guessability-blank.pdf",height=5,width=7.5)

d %>% filter(filename=="Korean_23_Motion_org") %>% View()

@@

### Figure: Guessability by study and iconicity rating

```{r figures_paper_guessability_rating}
tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
dp <- d
dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)

@@

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
ggsave("figures\\paper_panelAB_guessing_rating.png",height=5,width=9)

@@

### Figure: Guessability and cumulative iconicity

@@

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
ggsave("figures\\paper_panelAB.png",height=5,width=9)

@@

### Figure: Ratings and cumulative iconicity

@@

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
ggsave("figures\\paper_panelAB_ratings.png",height=5,width=9)

### Figure: Mappings and semantic domains

```{r figure_mappings_domains}

library(forcats)
dp.categories <- d %>%
  filter(category %notin% "Other") %>%
  mutate(category = fct_reorder(category, -C_cumulative, .fun='mean')) %>%
  mutate(category_binary = ifelse(category %in% c("Sound","Motion"),"SoundMotion","ColorvisualShapeTexture")) %>%
  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude) %>% # for control analysis
  mutate(C_strongestpredictors = C_aspect + C_length + C_modality ) # for another control analysis

dp.categories %>%
  group_by(category) %>%
  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

dp.categories %>%
  group_by(category) %>%
  summarise(n=n(),mean_C=mean.na(C_strongestpredictors),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

dp.categories %>%
  ungroup() %>%
  group_by(category_binary) %>%
  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_cumulative
group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_cumulative
t.test(group1,group2)
library(effsize)
cohen.d(group1,group2)

# also when we exclude C_modality? yes
dp.categories %>%
  group_by(category) %>%
  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)
dp.categories %>%
  group_by(category_binary) %>%
  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_nosound
group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_nosound
t.test(group1,group2)
cohen.d(group1,group2)

ggplot(data=dp.categories, aes(y=C_cumulative,x=category,colour=C_cumulative)) +
  theme_tufte() + theme(legend.position="none") +
  labs(title="Cumulative iconicity by semantic domain") +
  geom_half_violin() +
  geom_half_point() +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  NULL

ggplot(data=dp.categories, aes(y=C_cumulative,x=category,colour=C_cumulative)) +
  theme_tufte() + theme(legend.position="none") +
  labs(title="Cumulative iconicity by semantic domain") +
  geom_hex() +
  NULL

### Raincloud plots
While the icoplots provide a useful first visualisation, for in a paper we would probably want to use something like raincloud plots that show key features of the distribution across groups.

```{r visuals}

# panel A: scores by study and cumulative iconicity
ggplot(data=dp, aes(x=study,y=score,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank()) + 
  labs(title="Guessability by study and cumulative iconicity") +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white") +
  geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")

ggplot(data=dp, aes(x=C_cumulative,y=logodds)) +
  theme_tufte() +
  labs(title="Guessability (log odds) by cumulative iconicity",
       x="cumulative iconicity",
       y="guessability (log odds)") +
  geom_half_boxplot() +
  geom_half_point() +
  NULL

ggplot(data=dp, aes(x=C_any1,y=logodds)) +
  theme_tufte() +
  geom_half_boxplot() +
  geom_half_point() +
  NULL

ggplot(data=dp, aes(x=C_atleast2,y=logodds)) +
  theme_tufte() +
  geom_half_boxplot() +
  geom_half_point() +
  NULL

ggplot(data=dp, aes(x=C_ternary,y=logodds)) +
  theme_tufte() +
  geom_half_boxplot() +
  geom_half_point() +
  NULL

ggplot(data=dp, aes(x="all",y=score_z)) +
  theme_tufte() +
  geom_half_boxplot() +
  geom_half_point(side="l") +
  geom_half_violin(side="r") +
  NULL

ggplot(data=dp, aes(x=C_cumulative,y=logodds,color=C_cumulative,fill=C_cumulative)) +
  theme_tufte() +
  geom_half_violin() +
  geom_half_point_panel() +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9)

icoplot("C_cumulative")

ggplot(data=dp, aes(x=C_cumulative,y=logodds,color=C_cumulative,fill=C_cumulative)) +
  theme_tufte() +
  geom_half_violin() +
  geom_half_point_panel() +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  facet_grid(~ study)

### Dotplots

A quick look at some congruency measures. Uses a custom plotting function `icoplot()`, which is a wrapper for a ggplot object constructed with `geom_dotplot`.

```{r icoplots}

# all data points by language
icoplot() # icoplot() is loaded from TI_functions.R

icoplot("C_modality")
icoplot("C_aspect")
icoplot("C_length")
icoplot("C_magnitude")
icoplot("C_irregular")
icoplot("C_cumulative")

# how do the constituents of C_magnitude fare?
icoplot("C_magnitude")
icoplot("C_weight_voice")
icoplot("C_weight_vowel")
icoplot("C_weight_tone")

icoplot("C_any1")
icoplot("C_atleast2")
icoplot("C_ternary")
icoplot("C_simple")

# blank plot can be useful as background for slides
ggplot(data=dp, aes(x=study,y=score)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white",colour="#c9c9c9") +
  geom_dotplot(colour="white",fill="#c9c9c9",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")

# facet by language to discover crosslinguistic diversity
ggplot(data=dp, aes(x=study,y=score)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white",colour="#c9c9c9") +
  geom_dotplot(colour="white",fill="#c9c9c9",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")

@@

The simplest thing to do is a simple correlation. Looks like there is an r=0.30 correlation between logodds and C_cumulative.

@@

p0 <- cor.test(d$logodds,d$C_cumulative,method="pearson")$p.value
p1 <- cor.test(d.Collabra$logodds,d.Collabra$C_cumulative,method="pearson")$p.value
p2 <- cor.test(d.Language$logodds,d.Language$C_cumulative,method="pearson")$p.value
p.adjust(c(p0,p1,p2),method="bonferroni")

@@

probitlink(0.1823,inverse=T)-0.5
probitlink(-0.5242,inverse=T)-0.5

@@

probitlink(fixef(m8)[1],inverse=T)-0.5
probitlink(fixef(m8)[2],inverse=T)-0.5
probitlink(fixef(m8)[3],inverse=T)-0.5
probitlink(fixef(m8)[4],inverse=T)-0.5
probitlink(fixef(m8)[5],inverse=T)-0.5

@@

### Semantic domains
Our prediction is that Sound and Motion ideophones should show more structural correspondences than the other domains. This is borne out.

```{r coding_semanticdomains}

library(forcats)
dp.categories <- d %>%
  filter(category %notin% "Other") %>%
  mutate(category = fct_reorder(category, -C_cumulative, .fun='mean')) %>%
  mutate(category_binary = ifelse(category %in% c("Sound","Motion"),"SoundMotion","ColorvisualShapeTexture")) %>%
  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude) %>% # for control analysis
  mutate(C_strongestpredictors = C_aspect + C_length + C_modality ) # for another control analysis

dp.categories %>%
  group_by(category) %>%
  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

dp.categories %>%
  group_by(category) %>%
  summarise(n=n(),mean_C=mean.na(C_strongestpredictors),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

dp.categories %>%
  ungroup() %>%
  group_by(category_binary) %>%
  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_cumulative
group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_cumulative
t.test(group1,group2)
library(effsize)
cohen.d(group1,group2)

# also when we exclude C_modality? yes
dp.categories %>%
  group_by(category) %>%
  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)
dp.categories %>%
  group_by(category_binary) %>%
  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_nosound
group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_nosound
t.test(group1,group2)
cohen.d(group1,group2)

@@

We can also look in more detail at the relative differences by semantic domain by building a mixed model that predicts cumulative iconicity with domain as a predictor.

```{r coding_category_lme}

m0 <- lm(C_cumulative ~ category, data=d)
effsizes(m0)

@@

Another prediction is that Sound and Motion ideophones should receive higher iconicity ratings than other domains. This prediction is also borne out.

```{r rating_domains}

group1 <- subset(dp.categories, category_binary == "SoundMotion")$rating
group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$rating
t.test(group1,group2)
cohen.d(group1,group2)

@@

### Coding and ratings

As above, the simplest thing to do is run a simple correlation: there is an r=0.43 correlation between meaning iconicity ratings and cumulative iconicity.

```{r coding_ratings}

cor.test(d$rating_z,d$C_cumulative,method="pearson")

@@

As above, we can go a bit deeper by using mixed effects modelling. We expect a smaller effect of study on the ratings, as the difference in design in terms of selection of foils falls away. However, the studies are also different in terms of

@@

m0 <- lmer(rating ~ study + (1|ideophone), data=d)
m1 <- lmer(rating ~ study + C_cumulative + (1|ideophone), data=d)
anova(m0,m1)

@@

@@

### Control analysis: results stand even when leaving out C_modality

The `C_modality` congruence measure is intended to capture the fact that spoken words can show a range of acoustic similarities to sonic events, independently from other structural correspondences (in other words, it separates imagic iconicity from Gestalt iconicity and relative iconicity). We don't code for specific acoustic correspondences, because these would be by nature only apply to ideophones evocative of sound and would greatly complicate the coding scheme.

This makes it the least fine-grained coding category. And given that it applies to a semantic domain that is known, from prior work and first principles, to be easiest to guess, one could think of `C_modality` as just giving sound words a leg up in the cumulative scores. Therefore, as a control, it is useful to know whether the patterns we find hold up even if we disregard `C_modality` entirely.

```{r sound_off}

# we add C_nosound as a cumulative measure that excludes modality
d <- d %>%
  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude)

# recreate dp to include C_nosound
tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
dp <- d
dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)

d %>%
  group_by(C_nosound) %>%
  summarise(n=n(),score=mean.na(score))

ggplot(data=dp, aes(x=C_nosound,y=score_z,colour=C_nosound)) +
  theme_tufte() + theme(legend.position="none") +
  labs(title="Guessability (both studies) by cumulative iconicity",
       x="cumulative iconicity (excluding modality)",
       y=expression('guessability ('~italic(z)~')')) +
  geom_half_boxplot() +
  geom_half_point() +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  NULL

There is an r=0.26 correlation between logodds and C_sound across the studies. This is present in both studies.

```{r congruency_stats_nosound_1}

cor.test(d$logodds,d$C_nosound,method="pearson")

d.Collabra <- d %>% filter(study == "Collabra")
d.Language <- d %>% filter(study == "Language")
cor.test(d.Collabra$logodds,d.Collabra$C_nosound,method="pearson")
cor.test(d.Language$logodds,d.Language$C_nosound,method="pearson")

p0 <- cor.test(d$logodds,d$C_nosound,method="pearson")$p.value
p1 <- cor.test(d.Collabra$logodds,d.Collabra$C_nosound,method="pearson")$p.value
p2 <- cor.test(d.Language$logodds,d.Language$C_nosound,method="pearson")$p.value
p.adjust(c(p0,p1,p2),method="bonferroni")

The mixed effects models look much the same.
A null model `m0` predices logodds by paper, controlling for item-level differences. A model `m1` with cumulative congruence as an additional fixed effect fares significantly better.

For the fixed effects, m1 lists a 0.205 improvement in logodds for `C_nosound`. According to my interpretation this means that every added iconic feature leads to an ~8% boost in proportion correct. The lower baseline for the Language set is estimated as -0.513, which translates to a 19.5% guessability penalty.

```{r congruency_stats_nosound2}

m0 <- lmer(logodds ~ study + (1|ideophone), data=d)
m1 <- lmer(logodds ~ study + C_nosound + (1|ideophone), data=d)
anova(m0,m1)

probitlink(0.205,inverse=T)-0.5
probitlink(-0.513,inverse=T)-0.5

## Exploring coding sensitivity and congruency blind spots
What are we not capturing? Our coding and congruency measures are designed to be reproducible and human-codable. They are not exhaustive, and there may well be aspects that we are missing. Some ideophones may also be guessed relatively high for reasons other than form-meaning analogies (e.g., they had a very unattractive foil or presented unforeseen confounds). Here we explore the congruency data to find highly guessed ideophones that our coding did not (yet) identify as having iconic mappings. We also look at the other end: ideophones coded as having an iconic mapping but not doing well in the binary forced choice task.

If we sort by score most of the blindspots are from the Collabra study, which had a less forbidding design and therefore a higher baseline score.

```{r congruency_blindspots}

d %>%
  filter(C_cumulative == 0, score_z > 0.5) %>%
  dplyr::select(filename,ideophone,meaning,meaning_NL,language,logodds,score) %>%
  arrange(-score) %>% ungroup %>%
  slice(1:10)

Coding sensitivity varies by feature, in part because not all are equally frequent. For instance, `C_weight_tone` captures ideophones that have high tone/intonation and small/light meaning or conversely low tone/intonation and big/heavy meaning.

```{r coding_sensitivity}

icoplot("C_weight_tone")

d %>%
  group_by(language) %>%
  filter(C_weight_tone == 1) %>%
  summarise(n=n(),score=mean.na(score),rating=mean.na(rating))

d %>%
  group_by(language) %>%
  filter(C_weight_tone == 1) %>%
  View()

### Variation by language

Although we have focused on cross-linguistically attested features, features may also vary by language. For instance, irregularity may work somewhat in Japanese, Korean and Ewe, the 4 Semai ideophones scored as showing this mapping have a mean accuracy score of 0.45, dragging down the average. It looks like partial reduplication really has a different function in Semai, and one that is not captured by our coding scheme.

```{r coding_by_language}

d %>%
  group_by(language) %>%
  filter(C_irregular == 1) %>%
  summarise(n=n(),score=mean.na(score),rating=mean.na(rating),C=mean.na(C_cumulative))

d %>%
  group_by(language) %>%
  filter(C_irregular == 1) %>%
  dplyr::select(ideophone,meaning,language,category,score,rating,C_irregular,C_cumulative) %>%
  View()

mS0 <- lmer(logodds ~ study + (1|ideophone), data=d)
mS1 <- lmer(logodds ~ study + language + (1|ideophone), data=d) # fails to converge
mS2 <- lmer(logodds ~ study + C_aspect + C_modality + C_length + (1|ideophone), data=d)
mS3 <- lmer(logodds ~ study + C_aspect + C_modality + C_length + language + (1|ideophone), data=d)

anova(mS2,mS3)

fixef(mS3)

One might conclude from that language would be a relevant factor to take into account in the statistical models. But its influence is not very large. If we add it to the most complete model above, the fit is marginally improved, but the difference is not of the same magnitude as other factors. Also, language is partly confounded with quality of recordings, which are lab recordings for Ewe, Korean, Japanese and field recordings for Siwu and Semai. We can dummy code this as a new variable to check its effect.

```{r coding_by_audio}

d <- d %>%
  mutate(audio = ifelse(language %in% c("Siwu","Semai"),"field","lab"))

d %>%
  group_by(audio) %>%
  summarise(n=n(),score=mean.na(score),rating=mean.na(rating))

ggplot(data=d, aes(x=audio,y=score,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
  labs(title="Guessability by audio version",
       x="audio",
       y="guessability (% correct)") +
  geom_beeswarm(cex=2.5) +
  NULL

ggplot(data=d, aes(x=language,y=score,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
  labs(title="Guessability by audio version",
       x="language",
       y="guessability (% correct)") +
  geom_beeswarm(cex=2.5) +
  NULL

# ratings look like they have a distinctly non-normal distribution for field recordings, with a large bulge below 3 (= middle)
ggplot(data=d, aes(x=audio,y=rating,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
  labs(title="Iconicity ratings for 239 ideophones",
       y="mean iconicity rating") +
  geom_beeswarm(cex=2.5) +
  NULL

# and indeed the bulge isn't really there for any of the non-field languages (and arguably mostly attributable to the Semai recordings)
ggplot(data=d, aes(x=language,y=rating,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
  labs(title="Iconicity ratings for 239 ideophones",
       x="language",
       y="mean iconicity rating") +
  geom_beeswarm(cex=2.5) +
  NULL

```{r congruency_underperformers}
# the reverse is also interesting: which ideophones were coded as having iconic mappings yet were not guessed greatly?

d %>%
  filter(C_cumulative > 0, logodds < -0.25) %>%
  dplyr::select(filename,ideophone,meaning,meaning_NL,C_cumulative,language,logodds,score) %>%
  arrange(score) %>% ungroup() %>%
  slice(1:10)

>>>>

---
title: "Triangulating iconicity: Coding analysis"
author: "Mark Dingemanse, Stella Punselie, Bonnie McLean"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

@@

# get consensus coding data
d = read_excel("data/ideophones_coded.xlsx") %>% arrange(filename)

@@

d.scores = read_excel("data/ideophones_guessability.xlsx") %>%
  dplyr::select(-category)
d <- left_join(d,d.scores,by=c("ideophone","language","study" = "paper"))

@@

# get ratings data
d.ratings <- read_xlsx("data/ideophones_rated_means.xlsx") %>%
  dplyr::select(-category,-list,-item)
d <- left_join(d,d.ratings,by=c("filename","study","language")) |>
  mutate(filename = gsub("_org","",filename))

@@

ideophones_filename <- d |> ungroup() |> select(ideophone,filename) |>
  mutate(filename = str_to_lower(filename))

# load the full ratings data for some of the reporting in the paper
d.ratings.full <- read_xlsx("data\\ideophones_rated.xlsx") |>
  mutate(list = gsub(".*([A-Z])$", "\\1",item)) |>
  mutate(filename = gsub("^(.*)_.*$", "\\1", str_to_lower(item))) |>
  mutate(filename = gsub("Coll_","", filename)) |>
  select(-item)

# match the data to ideophones for easier lookup
d.ratings.full <- left_join(d.ratings.full,
                  ideophones_filename,by="filename",copy=TRUE)

@@

d %>%
  group_by(study) %>%
  dplyr::summarize(max=max(score),min=min(score),mean=mean.na(score))

@@

ggsave("figures\\fig2-guessability-blank.pdf",height=5,width=7.5)

@@

Let's plot the guessability by study and iconicity rating

```{r figures_paper_guessability_rating}

@@

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
#ggsave("figures\\paper_panelAB_guessing_rating.png",height=5,width=9)

@@

### Figure 3: Guessability and cumulative iconicity

@@

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
ggsave("figures\\fig3-panelAB.png",height=4,width=9,bg="white")

@@

### Figure 4: Ratings by study (A) and by semantic domain (B)

Panel figure for Study B providing a view of ratings per study and per semantic domain.

```{r fig4_panel}

#panel A: ratings by study
pA <- d |>  ggplot(aes(x=score_z,y=rating)) +   
  theme_tufte(base_size = 16) +
  theme(legend.position = c(0.2,0.9),
        legend.title = element_blank()) +
  scale_fill_manual(values=c("white","black")) +
  geom_point(aes(fill=study),
             position="jitter",
             shape=21,
             colour="black") +
  geom_smooth(method=loess,colour="black",alpha=0.5) +
  xlab("guessability (z)") + ylab("rating") +
  NULL

#panel B: ratings by semantic domain
pB <- d |>
  filter(category != "Other") |>
  ggplot(aes(x=reorder(category,-rating),y=rating,color=category)) +
  theme_tufte(base_size=16) +
  geom_half_boxplot(aes(middle=mean(rating)),show.legend=F) +
  geom_half_point(show.legend=F) +
  scale_colour_viridis_d(option="D",alpha=0.8) +
  xlab("category") + scale_x_discrete(labels = c("Sound","Motion","Shape", 
  "Texture","Colour/Visual"))

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.6,2))
ggsave("figures/fig4-panel_ratings.png",height=4,width=9,bg="white")

# a correlation test shows a Pearson correlation of .57 between iconicity rating
# and guessability
cor.test(d$rating_z,d$logodds,method="pearson")

Table 3: Examples of ideophones from the low/high rating/guessability quadrants

```{r example_ratings_guessables}

# these examples are given in Table 3 in the paper
examples <- c("ton ton","kpa",
              "slʔẽẽk","miɔmiɔ",
              "kodzokodzo","ɟtoonɟtoonɟtoon",
              "fututu", "plɒ̃s")
d |>
  filter(ideophone %in% examples) |>
  group_by(ideophone,language) |>
  dplyr::summarise(n=n(),
                   mean_rating = mean(rating),
                   score = score)

### Figure 5: Ratings and cumulative iconicity

@@

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
ggsave("figures/fig6-panelAB_ratings.png",height=5,width=9,bg="white")

@@

The simplest thing to do is a correlation.

@@

p0 <- cor.test(d$logodds,d$C_cumulative,method="pearson")$p.value
p1 <- cor.test(d.Collabra$logodds,d.Collabra$C_cumulative,method="pearson")$p.value
p2 <- cor.test(d.Language$logodds,d.Language$C_cumulative,method="pearson")$p.value
p.adjust(c(p0,p1,p2),method="bonferroni")
pvalues <-p.adjust(c(p0,p1,p2),method="bonferroni")

@@

# fixed effects
fixef(m1)

# turn this into a percentage under the simplifying assumption that the distance
# between iconicity measures is the same at every scale point
probitlink(fixef(m1)[[3]],inverse=T)-0.5
probitlink(fixef(m1)[[2]],inverse=T)-0.5

@@

probitlink(fixef(m8)[1],inverse=T)-0.5
probitlink(fixef(m8)[2],inverse=T)-0.5
probitlink(fixef(m8)[3],inverse=T)-0.5 # aspect
probitlink(fixef(m8)[4],inverse=T)-0.5 # modality
probitlink(fixef(m8)[5],inverse=T)-0.5 # length

@@

### Structural correspondences and semantic domains

Our prediction is that Sound and Motion ideophones should show more structural correspondences than the other domains. This is borne out.

```{r descriptive_semantic_domains}

dp.categories <- d %>%
  filter(category %notin% "Other") %>%
  mutate(category = forcats::fct_reorder(category, -C_cumulative, .fun='mean')) %>%
  mutate(category_binary = ifelse(category %in% c("Sound","Motion"),"SoundMotion","ColorvisualShapeTexture")) %>%
  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude) %>% # for control analysis
  mutate(C_strongestpredictors = C_aspect + C_length + C_modality ) # for another control analysis

# mean of C_cumuluative
dp.categories |>
  group_by(category) |>
  dplyr::summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

# mean of only the strongest predictors
dp.categories |>
  group_by(category) |>
  dplyr::summarise(n=n(),mean_strongest=mean.na(C_strongestpredictors),mean_rating=mean.na(rating),mean_score=mean.na(score)) |>
  arrange(-mean_strongest)

# binary comparison
dp.categories |>
  ungroup() |>
  group_by(category_binary) %>%
  dplyr::summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) |>
  arrange(-mean_C)

group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_cumulative
group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_cumulative
t.test(group1,group2)
effsize::cohen.d(group1,group2)

# also when we exclude C_modality? yes
dp.categories |>
  group_by(category) |>
  dplyr::summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) |>
  arrange(-mean_C)
dp.categories |>
  group_by(category_binary) |>
  dplyr::summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) |>
  arrange(-mean_C)

group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_nosound
group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_nosound
t.test(group1,group2)
effsize::cohen.d(group1,group2)

@@

### Study B: Rating iconicity

Here we look at the iconicity ratings for all ideophones across both studies.

```{r coding_ratings_descriptives}

d.ratings.full |>
  group_by(ideophone,language) |>
  dplyr::summarise(n=n(),mean=mean.na(rating),sd=sd.na(rating))

# get means and standard deviations across all ideophones and for each language and category
mean(d.allratings$rating)
  # 2.948339
sd(d.allratings$rating)

examples <- c("kpa","boo boo","zuratto", "plɒ̃s")
d.ratings.full |>
  filter(ideophone %in% examples) |>
  group_by(ideophone,language) |>
  dplyr::summarise(n=n(),
                   mean_rating = mean(rating),
                   sd_rating = sd(rating))

@@

There appears to be a positive correlation between the rating and guessing data.

### Ratings: numbers

Iconicity ratings by semantic domain

```{r ratings_by_domain}

d.ratings.full |>
  filter(category != "Other") |>
  group_by(category) |>
  dplyr::summarise(n=n(),
                   mean_rating=mean(rating),
                   sd_rating = sd(rating)) |>
  arrange(desc(mean_rating))

@@

```{r linear_regressions}

@@

How do iconicity ratings relate to structural correspondences?

A simple correlation test suggests a positive relation between iconicity rating and cumulative iconicity. As above, we can go a bit deeper by using mixed effects modelling. We expect a smaller effect of study on the ratings, as the difference in design in terms of selection of foils falls away.

@@

# simple correlation test
cor.test(d$rating_z,d$C_cumulative,method="pearson")

# lme
m0 <- lmer(rating ~ study + (1|ideophone), data=d)
m1 <- lmer(rating ~ study + C_cumulative + (1|ideophone), data=d)
anova(m0,m1)

@@

As above, we can assess the relative weight of different features using model comparison. We build models by adding the congruency measures in rank order of attestation in the coded data. Doing this we find that C_aspect, C_modality and C_length are responsible for the bulk of the variance. Compared to the null model that only represents the baseline difference between studies, model `m8` fares significantly better (log likelihood difference = 21.3).

@@
